Master thesis

### 2017-06-01 docs/master-thesis.tex
Disturbance v1 build

@@

@@

@@

\subsection{Outputs anlaysis ?}\label{outputs-anlaysis}

@@

\subsubsection{Resistance and resilience
metrics}\label{resistance-and-resilience-metrics-1}

\subsubsection{Biodiversity
partitioning}\label{biodiversity-partitioning-1}

>>>>

Most tropical timber production originates from selective logging, the
targeted harvesting of timber from species of interest. Consequently,
tropical sylviculture can be assimilated to a disturbance. The main
difference between a disturbance and selective logging is the targetting
of both species and individuals of interest. So we decided to first
asses unselective disturbance effect on tropical forest ecosystem to
subsequently better understand selective logging effect.

We first implemented a disturbance module inside TROLL model to simulate
unselective disturbance. 60 tree communities were then defined with
different levels of both specific and functional richness to explore the
role of biodiversity on tropical forest ecosystem answer to disturbance.
Next, different levels of disturbance were simulated on previous mature
communities. Finally, resilience of ecosystem major global variables
(related to carbon stock, forest dynamic and florisitic structure) from
simulations were partitioned between selection effects and niche
complementarity \citep{Loreau2001}.

@@

Disturbance module was designed in the simplest way in order to relate
the ecosystem answer to volume lost without any individuals nor species
targetting. For a given iteration \(disturb_{iter}\), individuals are
picked randomly with a uniform law on the number of trees. Selected
individuals are then removed without trigerring a treefall to avoid any
side effect. The operation is repeated untill the disturbance result in
a defined lost basal area (\(disturb_{intensity}\) in \% of BA).

@@

In order to assess the role of biodiversity in ecosystem answer to
disturbance, we needed to create a space of experiments encompassing
both variation of disturbance, biodiversity and time. Disturbance was
represented by percentage of basal area loss (0\%, 25\%, 50\% and 75\%).
Biodiversity was integrated with two components specific and functional
diversities. We used species richness \(SR\) to represents species
diversity (5, 25, and 125). Functional diversity can be related to
numerous components, and \citet{Borgy2017} argued for 5: richness,
divergence, regularity, overlap and mean. Because mature forest were
created from a bare soil with TROLL simulations, we could not control a
priori divergence, regularity and overlap but only assess theim before
diturbance. Consequently, we focused on functional richness with convex
hull volume \(CHV\) and functional mean with community weighted mean
\(CWM\). For each level of species richness \(SR\), we selected 20
communities with growing convex hull volume \(CHV\) but with a community
weighted means close to the regional species pool community weighted
means. Effectivelly, we did not wanted drastic change in community means
that could have more effect than functional richness itself. This design
of experiments resulted in 60 communities (\(5~SR*20~CHV\)) and 240
simulations (\(60 ~communities*4~levels~of~disturbance\)) over 600 years
(maturity being assumed after 500 years of regeneration \citep{Li}).
Figure \ref{fig:DOE} presents the design of experiment for communities
biodiversity after the forest mature were simulated, and thus before
disturbance. We obtained a broad range of both functionl dispersion
\(FDis\) and aboveground biomass \(AGB\) for simulated forest ecosystems
before disturbance.

@@

\subsection{Ecosystem answer anlaysis}\label{ecosystem-answer-anlaysis}

Tropical forest ecosystems provides numerous ecosystem services linked
to several ecosystem functions. We focused on few functions and related
metrics to analyse ecosystem answer to disturbance: carbon stock with
aboveground biomass (\(AGB\) in \(ton~C.ha^{-1}\)), forest dynamic with
number of stem above \(10~cm\) diameter at breast height (\(N10\)), and
floristic composition with \textbf{?} (\textbf{?} in \textbf{?}).

The resilience of metrics values post disturbance were assessed through
\citet{Henry2012} formula:

\begin{equation}
  R\left(t\right)=\frac{Recovery\left(t\right)}{Loss\left(t_d\right)} \approx \frac{X(t)}{X(t_d-1)}
  \label{eq:Resilience}
\end{equation}

The resilience of the system \(R(t)\) at the time \(t\) is described by
the ratio of recovery \(Recovery(t)\) at time \(t\) to loss suffered
\(Loss(t_d)\) at disturbance time \(t_d\). We transformed the equation
in the resilience of the system \(R(t)\) at the time \(t\) being the
ratio of the ecosystem metric \(X(t)\) at time \(t\) to the ecosystem
metric before the disturbance happened \(X(t_d-1)\) at time \(t_d-1\).
\(X(t_d-1)\) was calculated as the mean of the ecosystem output for 50
last years of the simulation of the mature forest (over 600 years).
Then, we calculated for each simulation the recovery time
\(t_{recovery}\) were the ecosystem reached back its stable state
defined as its state before disturbance. But some ecosystem metrics can
reach pre disturbance value wihtout revealing ecosystem recovery. For
instance, number of stem above \(10~cm\) diameter at breast height
\(N10\) will first decrease due to disturbance. Then \(N10\) will exceed
its initial value due to new seedlings recruitment before decreasing
towards its pre disturbanec value. Consequently, the first time \(N10\)
reaches back its pre disturbance value can not be considered as the
recovery time \(t_{recovery}\). In order to solve this issue, we
considered the ecosystem as stable again when its whole set of observed
metrics reach their pre disturbance values.

Biodiversity is not only a facet of the experimental design and an
ecosystem output through florisitic composition, but also interact on
ecosystem functioning and consequently on its answer to disturbance.
Biodiversity ecosystem functioning relation can be split in
complementarity and selection effect with \citet{Loreau2001}
partitioning:

\begin{equation}
  \begin{array}{c}
    NE = X_O - X_E = CE + SE \\
    CE = N* \overline{\Delta RX} \overline{M}\\
    SE = N*cov(\Delta RX,M)
  \end{array}
  \label{eq:BiodivPart}
\end{equation}

Biodiversity net effect \(NE\) is based on the difference between
ecosystem variable \(X\) observed value \(X_O\) within the community
mixture of species and its expected value \(X_E\) if species performance
were equal to their performance in monocultures. This effect can be
partitioned between complementarity effect \(CE\), representing niche
partitionning, positive interactions, and resource supply, and selectvie
effect \(SE\) due to dominant species pool driving the ecosystem. Both
metrics depend on the variation of relative ecosystem variable
\(\Delta RX\):

\begin{equation}
  \Delta RX_{sp} = \frac{X_{sp}(mixture)}{X_{sp}(monoculture)} - P_{sp}
  \label{eq:DeltaRY}
\end{equation}

\(X_sp\) is the ecosystem variable value for one species either in
mixture \(X_{sp}(mixture)\) or in monoculture \(X_{sp}(monoculture)\).
\(P_{sp}\) is the proportion of the species in the mixture represented
by species relative abundance. Consequently, \(CE\) averages diversity
effects of all species presents in the mixture (both negatives and
positives). Whereas \(SE\) become positive when dominant species
outperform theimselves in mixture than in monoculture, and negative when
less dominant species outperform theimselves in mixture than in
monoculture \citep{Tobner2016}.

Recovery trajectories of ecosystem variable after disturbance were
partitioned between complementarity effect \(CE\) and selection effect
\(SE\). In order to do that, the design of experiment was repeated for
each species indivdually representing 652 simulations of monoculture.

@@
